---
title: "Databrary 2.0 schema development"
output: 
  html_notebook:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
params:
  db_login: your@email.com
---

```{r set-up, include=FALSE}
library(tidyverse)
```

# Purpose

To document Rick Gilmore's work on the Databrary 2.0 schema.

# Resources

Jeff Spies, Swapnaa Jayaraman, and Rick Gilmore are also using this Google Doc as a working document:

<https://docs.google.com/document/d/1_becXD6YX6lqgn4LK0IhgWII5C_pBzcQWFGDLx-SY_Y/edit?usp=sharing>

## Sites

- <https://re3data.org>
- DataCite: <https://www.datacite.org>
    - DataCite Metadata Schema v4.3 Properties Overview. <https://support.datacite.org/docs/schema-properties-overview-v43>

I'm using DB1 to refer to Databrary 1.0 and DB2 to refer to Databrary 2.0.

# Type by type

Login to Databrary by running `rmarkdown::render("schema-development.Rmd", params = list(db_login="YOUREMAIL@PROVIDER.COM"))` from the RStudio console.

```{r}
databraryapi::login_db(params$db_login)
```

# Repository-wide

The re3data.org specification adds some metadata elements that may be helpful to support.

## Institutions

### DB1

In DB1, all Institutions are instances of `party`, but with some special properties.

The DB1 API returns the following data for an Institution like NYU.

```{r}
databraryapi::download_party(8)
```

### Schema suggestions

Based on the CollegeOrUniversity schema (http://schema.org/CollegeOrUniversity), I suggest the following mappings:

| DB1           | Suggested mapping | Schema URL | Comments |
|---------------|-------------------|------------|----------|
| `id`          | `identifier`      | http://schema.org/identifier  | Use structured scheme for institutional IDs |
| `sortname`    | `alternateName`   | http://schema.org/alternateName | Add official name as separate entity |                         
| `url`         | `url`             | http://schema.org/url           | |
| `institution` |                   |                                 | Boolean value in DB1; needed in DB2? |
| https://nyu.databrary.org/party/8/avatar?size=500 | `logo` | http://schema.org/logo | |

I also suggest the following additions:

| DB1           | Suggested mapping | Schema URL | Comments |
|---------------|-------------------|------------|----------|
|               | `name`            | http://schema.org/name | New field with official name, here New York University |
|               | `address`         | http://schema.org/address | |
|               | `contactPoint`    | http://schema.org/contactPoint  |  |

## Users

In DB1, all users are instances of `party`. For DB2, we are considering creating a special type of `Person` called an `AccountHolder`.

### DB1

Here, for example, is what the DB1 API returns for Rick Gilmore:

```{r}
databraryapi::download_party(6)
```

These are largely self-explanatory except for `permission`. 
Let's investigate this.

```{r}
db_const <- databraryapi::assign_constants()
str(db_const)
```

It's pretty straightforward.
The `permission` is one of `r db_const$permission`. I have `r db_const$permission[5]` because it's my profile, possibly.

Yes, that's right.
See what happens when I query Karen's:

```{r}
databraryapi::download_party(5)
```

There is no `permission` defined.
I'm not sure how we'll map this in the new schema.

### Schema suggestions

| DB1           | Suggested mapping | Schema URL | Comments |
|---------------|-------------------|------------|----------|
| `id`          | `identifier`      | http://schema.org/identifier | Use structured scheme for IDs of people, maybe prefix current party ID with `p` |
| `sortname`    | `familyName`   | http://schema.org/familyName |  | 
| `prename`     | `givenName`    | http://schema.org/givenName  | |
| `orcid`       |                |                              | |
| `affiliation` |  `worksFor`     |  http://schema.org/worksFor  | |
| `url`         | `url`             | http://schema.org/url           | |
| `email`       | `email`           | http://schema.org/email         | |
| `permission`  |                   |                                 | |
| https://nyu.databrary.org/party/6/avatar?size=500 | `image` | http://schema.org/logo | |

With the following addition(s):

| DB1           | Suggested mapping | Schema URL | Comments |
|---------------|-------------------|------------|----------|
|               | `address`         | http://schema.org/address | |
|               | `additionalName`   | http://schema.org/additionalName | Middle name |
|               | `honorificPrefix`  | http://schema.org/honorificPrefix | e.g., Dr. |
|               | `honorificSuffix`  | http://schema.org/honorificSuffix | e.g., Ph.D. |
|               | `jobTitle`         | http://schema.org/jobTitle | |
|               | `telephone`        | http://schema.org/telephone | |

## Volumes

In DB1, we use the term volume, and so does the API. We are considering using the term `Project` in DB2.

### DB1

Here is the structure of the output from the DB1 API for a volume, in this case, the default is volume 1 (https://nyu.databrary.org/volume/1):

```{r}
str(databraryapi::download_containers_records())
```

We also have the project's icon at <https://nyu.databrary.org/volume/1/thumb?size>, so apparently the `thumb` variable holds the icon.

#### Questions

1. How is the date added information represented?
2. How are external links represented?
3. Is the HOW TO CITE generated from a function or is it represented separately as data?

### Schema suggestions

Let's try the DataCatalog (http://schema.org/DataCatalog) schema to start mapping the volume-level items.

| DB1             | Suggested mapping | Schema URL | Comments |
|------------------|-------------------|------------|----------|
| `id`             | `identifier`      | http://schema.org/identifier | |
| `name`           | `name`            | http://schema.org/name       | |
| `body`           | `abstract`        | http://schema.org/abstract   | |
| `doi`            | `doi`             | from re3data.org (property ID 1.2) |  |
| `creation`       | `dateCreated`     | http://schema.org/dateCreated ||
| `owners`         | `creator`         | http://schema.org/creator     | How to handle different types of 'ownership' |
| `permission`     |                   |                               | See related discussion above |
| `publicshareful` |                   |                               ||
| `publicaccess`   |                   |                               ||
| https://nyu.databrary.org/volume/1/thumb?size | `thumbnailUrl` | http://schema.org/thumbnailUrl | |          

We'll deal with the `containers` array separately, as that is where the spreadsheet/session data reside.

The following seem like logical additions to the volume level schema:

| DB1           | Suggested mapping | Schema URL | Comments |
|---------------|-------------------|------------|----------|
|               | `dateModified`    | http://schema.org/dateModified | |
|               | `datePublished`   | http://schema.org/datePublished | |
|               | `dataset`         | http://schema.org/dataset | |
|               | `measurementTechnique` | http://pending.schema.org/measurementTechnique | |
|               | `citation`        | http://schema.org/citation | But see `doi` query |
|               | `conditionsOfAccess` | http://pending.schema.org/conditionsOfAccess | |
|               | `contributor`       | http://schema.org/contributor | Possibly use for non-authorial/citation contributors |
|               | `copyrightHolder`    | http://schema.org/copyrightHolder | |
|               |                      | http://schema.org/copyrightYear | |
|               | `creativeWorkStatus` | http://pending.schema.org/creativeWorkStatus | |
|               | `funder`             | http://schema.org/funder | Not sure how funders are handled now |
|               | `hasPart`            | http://schema.org/hasPart | For subcomponents |
|               | `inLanguage`         | http://schema.org/inLanguage | |
|               | `interactionStatistic` | http://schema.org/interactionStatistic |  For view/download counts |
|               | `isPartOf`           | http://schema.org/isPartOf | For relations between projects |
|               | `keywords`           | http://schema.org/keywords | |
|               | `publication`         | http://schema.org/publication | |
|               | `version`             | http://schema.org/version | |
|               | `sameAs`              | http://schema.org/sameAs | |

### DataCite recommendations

The DataCite Metadata Schema v4.3 Properties Overview. <https://support.datacite.org/docs/schema-properties-overview-v43> makes specific recommendations about what properties are mandatory, which are recommended, and which are optional.

| ID | Property	| Obligation | Comments |
|----|----------|------------|----------|
| 1 | Identifier (with mandatory type sub-property) | M | Maps to `identifier` |
| 2 | Creator (with optional name identifier and affiliation sub-properties) | M | Maps to `creator` |
| 3 | Title (with optional type sub-properties) | M | Maps to `name` |
| 4 | Publisher | M | `publisher` in schema.org |
| 5 | PublicationYear | M | `datePublished` in schema.org |
| 10 | ResourceType (with mandatory general type description sub-property) | M | `additionalType` from schema.org? |

It seems prudent to try to include the mandatory DataCite properties.

The recommended (R) and optional (O) properties are as follows:

| ID | Property	                 | Obligation | Comments |
|----|---------------------------|------------|----------|
| 6  | Subject (with scheme sub-property) | R | | 
| 7  | Contributor (with type, name identifier, and affiliation sub-properties) | R | `contributor` |
| 8  | Date (with type sub-property)      | R | Use more specific `dateModified`, etc.?|
| 9  | Language                           | O | `inLanguage` |
| 11 | AlternateIdentifier (with type sub-property) | O | |
| 12 | RelatedIdentifier (with type and relation type sub-properties) | R | |
| 13 | Size                               | O | |
| 14 | Format                             | O | Relevant since Projects are heterogeneous? |
| 15 | Version                            | O | `version` in schema.org |
| 16 | Rights                             | O | Support with some version of Databrary access rights? |
| 17 | Description (with type sub-property) | R | What is relationship to `abstract`? |
| 18 | GeoLocation (with point, box and polygon sub-properties) | R | Derived from individual components of collection? |
| 19 | FundingReference (with name, identifier, and award related sub-properties) | O | Extension of `funder`? |

Here is a sample XML file for a simple dataset: <http://schema.datacite.org/meta/kernel-4.3/example/datacite-example-dataset-v4.xml>


## Sessions

In DB1, we show collections of data in a spreadsheet-like interface. 
A collection of data involving people observed on a specific date and time in a particular location doing similar things is called a *session*.
The DB1 API indexes sessions as `slot`s.
Each `slot` is given a unique integer ID.
For example, <https://nyu.databrary.org/volume/1/slot/25952/-> links to `slot` 25952 in volume 1, the volume that contains information about Databrary's board meetings.

### DB1

The API returns a table called `containers`:

```{r}
vol1 <- databraryapi::download_containers_records()
str(vol1)
```

The `containers` table contains the following fields.

```{r}
vol1_cont <- vol1$containers
names(vol1_cont)
```

For volume 1, this is an array of with dimensions `r dim(vol1_cont)`, one row for each of the "folders" or slots/sessions in the volume.
The `id` variable holds the `slot` index. 

The number of variables collected for each session/`slot` varies across studies.
Consider one of the PLAY datasets, for example.

```{r}
vol899 <- databraryapi::download_containers_records(899)
vol899_containers <- vol899$containers
names(vol899_containers)
```

The variables for the session/`slot` are identical to those for volume 1.

However, the data returned in the spreadsheet about the participants are stored in a `records` table with three variables:

```{r}
vol899_records <- vol899$records
vol899_records
```

The `measures` variable is a data.frame.
The `category` variable refers to the type of information recorded in the spreadsheet:

```{r}
db_const <- databraryapi::assign_constants()
str(db_const$metric)
```

The mappings are stored in a site-wide array af constants I access using `databraryapi::assign_constants`.
This is a well-structured array that we can probably map to open schema.

### Questions

1. What do the `id` and `category` variables within the `records` table refer to?

The ID appears to refer to these session/`slot` data, independent of the `measures`. I believe this is because it's possible that there could be more than one person recorded in a session/`slot`.
So, there must be a way to link multiple `records$id`s to a `containers$id`.

The `category` refers to the same grouping of participant-level variables as described below.

2. What do the `category` and `release` variables within the `metrics` do?

Based on the following...

```{r}
str(db_const$category)
```

It looks like the `category` refers to an aggregation of data about the session.
The relevant values are in \{'participant', 'pilot', 'exclusion', 'condition', 'group', 'task', 'context'\}.

Similarly, `release` is one of the following values:

```{r}
db_const$release
```

Since most of the data items contain a value of 3, except for `birthdate` and `disability`, I'm guessing that this is used to filter these variables from being shown when the study is shared.
